---
title: "2022-NW-Trawl"
author: "Tim van der Stap"
date: "6/16/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(readxl)
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

```{r data download, include=FALSE}
download.file("https://github.com/international-year-of-the-salmon/2022-NW-Data-Template/blob/main/IYS2022_NW_Data.xlsx?raw=true", here("IYS_data_template", "IYS2022_NW_Data.xlsx"))
event <- read_excel("./IYS_data_template/IYS2022_NW_Data.xlsx", sheet = "Sampling_Event_Info")
```
event <- read_excel("IYS_data_template/IYS2022_TINRO.xlsx", 
    sheet = "4. SAMPLING EVENT INFO") %>% 
  filter(Event_Type == "CTD") %>% 
  mutate(Time_Start = str_sub(Time_Start, -8, -1), # Remove leading date
         Time_End = str_sub(Time_End, -8, -1),
         Date_Time = paste0(format_ISO8601(as.POSIXct(paste0(as_date(
           paste0(Year, "-", Month, "-", Day)), " ", Time_Start), tz="UTC")), "Z"),
        # convert numeric fractional of day to minutes
         Sampling_Duration_Minutes = round(Sampling_Duration_Minutes * 1440, 0)) %>%
  select(-Day_Night, -Tow_speed_kilometers_per_hour, -Tow_distance_nautical_miles, -Swell_height_Meters, -Weather_description, -Comments) 
